Main Analyses of the Combined Dataset

### 2021-05-13 manuscript/manuscript.tex
table and biomass sections

The approximate census sizes I have estimated here

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/species_counts.pdf}

  \label{suppfig:species_counts}
\end{figure}

\begin{table}[ht]
  \resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{lr|rr|rrrr|r}
  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{4}{c}{\bf Present study}  &  \\
  phylum & total species ($T$) & biomass ($B$) & prop. biomass & biomass ($b$) &  prop. biomass & num. species ($n$) &prop. total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
  \hline
  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.78 \times 10^{-4}$ & 0.0100 &  68 & $5.41 \times 10^{-5}$ & 4.29 \\ 
  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.70 \times 10^{-2}$ & 0.9719 &  68 & $1.26 \times 10^{-3}$ & 24.70 \\ 
  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & $1.76 \times 10^{-4}$ & 0.35 \\ 
  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.54 \times 10^{-4}$ & 0.0163 &  13 & $1.36 \times 10^{-4}$ & 16.60 \\ 
  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.14 \times 10^{-5}$ & 0.0011 &   2 & $1.25 \times 10^{-4}$ & 2.51 \\ 
  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $3.89 \times 10^{-6}$ & 0.0001 &   1 & $4.00 \times 10^{-5}$ & 4.87 \\ 
\end{tabular}}

  \caption{How the total carbon biomass estimates by phylum from
    \textcite{Bar-On2018-kc} compare to the implied biomass estimates from this
    study. All biomass estimates are carbon biomass, and the proportions are of
    total biomass with respect to the study. The proportion of biomass in this
    study compared to the Bar-On et al. estimates \textcite{Bar-On2018-kc}
    indicates chordates are over-represented and arthropods are
    under-represented in the present study. Total species by phylum estimates
  are from
\textcite{Reaka-Kudla1996-fe,Nicol1969-qx,Zhang2013-zc,Chapman2009-fw}. The
ratio column is the ratio of total biomass implied by the $N_c$ estimates of
each species in a phylum to the actual biomass of that phylum.}

\end{table}

>>>>

I validated the approximate census sizes I have estimated here using a few
approaches. First, to check broad consistency, I compared the implied biomass
from my estimates of body mass and census size per species with estimates of
the total carbon biomass on earth by phylum \parencite{Bar-On2018-kc}. For
species $i$ with wet body mass $m_i$ and census size $N_i$, the implied biomass
is $m_i N_i$. For all species in a phylum $S$, this total sample biomass is
$b_S = \sum_{i \in S} m_i N_i$. I then compare this wet biomass to the carbon
biomasses by phylum by \textcite{Bar-On2018-kc}. Across animal species, the
ratio of dry to wet body mass, and carbon body mass dry body mass varies
little. In their study, Bar-On et al. assume wet body mass has a 70\% water
content, and 50\% of dry body mass is carbon mass, leading to a wet body mass
to carbon mass factor of $\nicefrac{1-0.7}{0.5} = 0.15$. I use this factor to
convert the total wet biomass to carbon biomass per phylum.

\begin{table}[ht]
  \resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{lr|rr|rrrrr|r}
  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{5}{c}{\bf Present study}  &  \\
  phylum & total species ($T$) & biomass ($B$) & prop. biomass & biomass ($b$) &  prop. biomass & num. species ($n$) & factor overrepresented &prop. total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
  \hline
  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.78 \times 10^{-4}$ & 0.0100 &  68 & 0.02 & $5.41 \times 10^{-5}$ & 4.29 \\
  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.70 \times 10^{-2}$ & 0.9719 &  68 & 2.90 & $1.26 \times 10^{-3}$ & 24.70 \\
  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & 0.01 & $1.76 \times 10^{-4}$ & 0.35 \\
  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.54 \times 10^{-4}$ & 0.0163 &  13 & 0.21 & $1.36 \times 10^{-4}$ & 16.60 \\
  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.14 \times 10^{-5}$ & 0.0011 &   2 & 0.03 & $1.25 \times 10^{-4}$ & 2.51 \\
  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $3.89 \times 10^{-6}$ & 0.0001 &   1 & 0.02 & $4.00 \times 10^{-5}$ & 4.87 \\
\end{tabular}}
  \caption{How the total carbon biomass estimates by phylum from
    \textcite{Bar-On2018-kc} compare to the implied biomass estimates from this
    study. All biomass estimates are carbon biomass, and the proportions are of
    total biomass with respect to the study. The proportion of biomass in this
    study compared to the Bar-On et al. estimates \textcite{Bar-On2018-kc}
    indicates chordates are overrepresented and arthropods are underrepresented
    in the present study; the factor that each phylum is overrepresented is
    given in the eighth column. Total species by phylum estimates are from
    \textcite{Reaka-Kudla1996-fe,Nicol1969-qx,Zhang2013-zc,Chapman2009-fw}. The
    ratio column is the ratio of total biomass implied by the $N_c$ estimates
    of each species in a phylum to the actual biomass of that phylum.}
    \label{supp:biomass-tbl}
\end{table}

First, I compared the relative carbon biomass in this study to the relative
carbon biomass on earth per phylum. This shows that this study's sample is over
represents chordate biomass (by a factor of $\approx 3$), and under represents
in arthropod biomass (by a factor of $0.02$) relative to the proportion of
carbon biomass of these phyla on earth (see column eight of Supplementary Table
\ref{supp:biomass-tbl}. Second, to check whether the carbon biomass per phylum
in the sample was broadly consistent with the total on earth by phylum ($B_S$
for phylum $S$), I calculated the expected sample biomass if species were
sampled randomly from the total species in a phylum, ($B_S \times
\nicefrac{n_S}{T_S}$, where $n_S$ is the total number of species in the sample
in phylum $S$, $T_S$ is the total number of species in phylum $S$ on earth).
The fraction of total species on earth included in the sample in this study is
depicted in Supplementary Figure \ref{suppfig:species_counts}.

Next, I look at the ratio of sample biomass per phylum, $b_S$ to this expected
biomass per phylum (Supplementary Table \ref{supp:biomass-tbl}). The
consistency is quite close for this rough approach and the non-random sample of
taxa included in this study. The carbon biomass estimates for chordates implied
by the census size estimates are $\approx 25$-fold higher than expected, but is
well within reasonable expectations given that the chordate sample includes
many larger-bodied domesticated species (and is a biased sample in other ways).
Similarly, the implied arthropod carbon biomass is quite close to what one
would expect. Overall, these values indicate that the census size estimates
here do not lead to implied biomasses per phylum that are outside the range of
plausibility.

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/species_counts.pdf}
  \caption{The fraction of total species on earth included in this study's
  sample, per class. The color of the points represents phylum, and the size of
the point represents the absolute number of species by class.}
  \label{suppfig:species_counts}
\end{figure}

### 2021-05-13 notebooks/main_analysis.Rmd
table and biomass sections

bm_tbl <- biomass %>% mutate(total_species = as.integer(total_species)) %>%
           select(phylum, total_species, bd_mass, bd_prop_mass, 
                  total_carbon_mass, prop_carbon_mass, nspecies, frac, factor)

print(xtable(bm_tbl, digits=c(1, 2, -2, 2, 4, -2, 4, 2, -2, 2)), 
      include.rownames=FALSE,
      tabular.environment='tabular',
      math.style.exponents=TRUE, type='latex')

>>>>

bm_tbl <- biomass %>% mutate(total_species = as.integer(total_species)) %>%
           mutate(over = prop_carbon_mass / bd_prop_mass) %>%
           select(phylum, total_species, bd_mass, bd_prop_mass, 
                  total_carbon_mass, prop_carbon_mass, nspecies, over, frac, factor)

print(xtable(bm_tbl, digits=c(1, 2, -2, 2, 4, -2, 4, 2, 2, -2, 2)), 
      include.rownames=FALSE,
      tabular.environment='tabular',
      math.style.exponents=TRUE, type='latex')
